---
title: "2016 pif3gi2 timelapse"
author: "Kazu Nozue"
date: "October_1, 2016"
output: html_document
---
# pif3 gi2 hypocotyl timelapse growth analyis
## written by Kazunari Nozue (Maloof lab)
## collaboration with Maria (Kay lab)
### history
* for exp1 and exp2 (different start day and timing of image capture) (100316)
### outline

### prerequisite
```{r}
library(ggplot2)
library(dplyr)
# functions used in this script
## vlookup
#Vlookup in R (https://gist.github.com/jnmaloof/7367450)
#Version 0.3 November 12, 2013
#Return senesical results if return column is a factor
#Version 0.2 November 11, 2013
#Require first column of table to be numeric if range lookup is being done
#Change defaults to larger=FALSE
#Julin Maloof
vlookup <- function(ref, #the value or values that you want to look for
                    table, #the table where you want to look for it; will look in first column
                    column, #the column that you want the return data to come from,
                    range=FALSE, #if there is not an exact match, return the closest?
                    larger=FALSE) #if doing a range lookup, should the smaller or larger key be used?)
{
  if(!is.numeric(column) & !column %in% colnames(table)) {
    stop(paste("can't find column",column,"in table"))
  }
  if(range) {
    if(!is.numeric(table[,1])) {
      stop(paste("The first column of table must be numeric when using range lookup"))
    }
    table <- table[order(table[,1]),] 
    index <- findInterval(ref,table[,1])
    if(larger) {
      index <- ifelse(ref %in% table[,1],index,index+1)
    }
    output <- table[index,column]
    output[!index <= dim(table)[1]] <- NA
    
  } else {
    output <- table[match(ref,table[,1]),column]
    output[!ref %in% table[,1]] <- NA #not needed?
  }
  dim(output) <- dim(ref)
  output
}
```
# data formating
```{r}
# setup working directory
if(Sys.info()["user"]=="nozue") {
  homedir<-"/Volumes/Data6/data_JM4/2016pifgi2_timelapse"
} # customize homedir
if(Sys.info()["user"]=="jmaloof") {
  homedir<-"~/git/Marian_pif3gi2_timelapse/"
} # customize homedir
setwd(homedir)
# 
# reading file time stamp
## This file was created in MSDOS (Celia cam computer) by "dir" command
filetime<-read.table("2016pif3gi2filetime.txt",skip=7,nrows=869)
head(filetime)
tail(filetime)

# reading file name used in exp1
exp1.filename<-read.table("exp1.file.txt",skip=1)
# add 336 to 1 (frmae number in exp1 avi )
exp1.filename$frame<-1:nrow(exp1.filename)
# reading file name used in exp2
exp2.filename<-read.table("exp2.file.txt",nrow=533) # skip the last line "exp2.file.txt"
#exp2.filename$frame<-1:nrow(exp2.filename) # is this true? no 
exp2.filename$frame<- c(((533-309)+1):533, (310-309):(533-309))# is this true? OK. Probably I should fix it becasue this is dirty. (100316)

# reading hypocotyl length data
hyp.data<-read.csv("Marian_gi2pif3_hyp_data.csv")
# need to clean up data (no duplicate)
hyp.data$id<-paste(sub("(2016pif3gi2exp)(1A_|1B_|2A_|2B_)(pif3gi2|pif3|gi2|Col)(.avi:)([[:digit:]]+)([[:print:]]+)","\\1\\2\\3\\4\\5",hyp.data$label),hyp.data$rep,sep="_")
hyp.data[duplicated(hyp.data$id),] # OK
# change pixel into mm
scale<- 95/1220.3213 # 95 mm is 1220.3213 pixel
hyp.data$mm<-hyp.data$pixel * scale


# merge time stamp data and hyp.data
## extract avi movie frame name
hyp.data$frame<-as.integer(sub("(2016pif3gi2exp)(1A_|1B_|2A_|2B_)(pif3gi2|pif3|gi2|Col)(.avi:)([[:digit:]]+)([[:print:]]+)","\\5",hyp.data$label)) 

# obtaining file date/time (exp1)
hyp.data[hyp.data$exp==1,"file"]<-as.character(vlookup(hyp.data[hyp.data$exp==1,"frame"],exp1.filename[,2:1],2))
str(hyp.data)
# obtaining file date/time (exp2)
hyp.data[hyp.data$exp==2,"file"]<-as.character(vlookup(hyp.data[hyp.data$exp==2,"frame"],exp2.filename[,2:1],2))
## obtain file date/time stamp (exp1 and exp2)
hyp.data$date<-vlookup(hyp.data$file,filetime[,c("V4","V1")],2)
hyp.data$time<-vlookup(hyp.data$file,filetime[,c("V4","V2")],2)
## converted to "POSIXlt" (see ?strptime)
hyp.data$date.time<-strptime(paste(hyp.data$date,hyp.data$time,sep="/"),"%m/%d/%Y/%H:%M")

## June 27 is day3 after stratification (start day for exp1 is 
exp1.start.date.time<-strptime("6/24/2016/7:00","%m/%d/%Y/%H:%M") # start date time for exp1
exp2.start.date.time<-strptime("7/1/2016/7:00","%m/%d/%Y/%H:%M") # start date time for exp2

## day-night cycle (7:00 - 15:00 is day)
# calculate hours after stratification 
# cf. hyp.data$date.time - start.date.time gave me units="days"
# for exp1
hyp.data[hyp.data$exp==1,"after.stratification_hr"]<-
  as.numeric(gsub(" hours","",difftime(hyp.data[hyp.data$exp==1,"date.time"],exp1.start.date.time,units="hours"))) # see more in ?difftime
# for exp2
hyp.data[hyp.data$exp==2,"after.stratification_hr"]<-
  as.numeric(gsub(" hours","",difftime(hyp.data[hyp.data$exp==2,"date.time"],exp2.start.date.time,units="hours"))) # see more in ?difftime
# 
str(hyp.data)
head(hyp.data)
summary(hyp.data)

```

# calculate growth rate. how to ? exp1 and exp2 has different. And then smoothing curve as Julin did in Holtan (2011) Plant Physiol?
```{r}
# from Julin's scripts "TLsmoothMND MAR.R"
calcRate <- function(x) { #no running average
  x <- x[order(x$after.stratification_hr),] 
  x$rate <- NA
  for (i in 2:nrow(x)) {
    x$rate[i] <- (x$y[i]-x$y[i-1]) / (x$after.stratification_hr[i] -x$after.stratification_hr[i-1])
  }
  x
} 

names(hyp.data)<-sub("mm","y",names(hyp.data))
hyp.data$sample<-with(hyp.data,paste(exp,plate,genotype,rep,sep="_")) 
rate.list <- by(hyp.data,hyp.data$sample,calcRate) 
tl.rates <- rate.list[[1]]              #need to recompile the list of dataframes into one dataframe.  this will hold it

for (i in 2:length(rate.list)) tl.rates <- rbind(tl.rates,rate.list[[i]]) #merge into one frame

tl.rates3<-tl.rates[!is.na(tl.rates$rate),]
```

## Calculate sample size
```{r}
samples <- unique(hyp.data$sample)
genotypes <- as.character(unique(hyp.data$genotype))
sample.numbers <- sapply(genotypes,function(g) {
  sum(grepl(g,samples))
})
sample.numbers
```


- Ideally, names of the different lines should be changed to be consistent with the ones use across the manuscript (col0, gi-2, pif3-1, gi-2;pif3-1).
-  Axis text fonts should be increased to be legible.
- Ideally (but only if possible), I’d like the colors to also match the ones used across the ms (col0 black, gi-2 orange, pif3-1 yellow, gi-2;pif3-1 green).

# plot data

## height

```{r}
# plot (height version)
p.height<-ggplot(hyp.data,aes(x=after.stratification_hr,y=y,color=as.factor(rep),shape=factor(exp))) + geom_point()+facet_grid(genotype~exp)
p.height

```


## rates
```{r,fig.height=3,fig.width=4}

labels <- c("Col"="col0", "gi2"="gi-2", "pif3"="pif3-1", "pif3 gi2" = "gi2;pif3-1")
colors <- c("Col"="black", "gi2"="orange", "pif3"="yellow", "pif3 gi2" = "green")


night <- seq(8,max(tl.rates$after.stratification_hr),24)
day <- seq(24,max(tl.rates$after.stratification_hr),24)
if (length(night)>length(day)) day <- c(day,max(tl.rates$after.stratification_hr))
bottom <- rep(-0.02,length(night))
top <- rep(0.11,length(night)) # for mm

#data frame to hold coordinates for nighttime rectangles
rectangles <- data.frame(night,day,top,bottom,rate=0,after.stratification_hr=0)
rm(night,day,top,bottom)

p.growthrate<- ggplot(data=tl.rates3,aes(y=rate,x=after.stratification_hr))  
p.growthrate <- p.growthrate + facet_grid(genotype~., labeller = labeller(genotype=labels))
p.growthrate <- p.growthrate + geom_rect(data=rectangles,mapping=aes(xmin=night,xmax=day,ymin=bottom,ymax=top),fill="grey80")
p.growthrate <- p.growthrate + geom_smooth(aes(color=genotype),stat="smooth",fill="grey20",method="loess",span=.1,n=200) # needs to play with span and n
p.growthrate <- p.growthrate + theme(panel.background = element_rect(fill = "white")) + 
  labs(x="Time (hr)",y="Growth rate (mm/hr)") +
   theme(axis.title = element_text(face="bold"))
p.growthrate <- p.growthrate + coord_cartesian(xlim=100:max(tl.rates3$after.stratification_hr),ylim=c(-0.015,0.1))
p.growthrate <- p.growthrate + scale_color_manual(values = colors, guide=FALSE) + scale_y_continuous(breaks=c(0,.05,.1))
p.growthrate

ggsave(p.growthrate,file="Maria_pif3gi2_timelapse_v2.png",width=4,height=3) 
# Time (h): time after stratification (hr) (cf. Nozue (2007) use "zero indicates dawn of the fourth day")
#
```

# plot data on single facet
```{r,fig.height=2.7,fig.width=3.3}

p.growthrate<- ggplot(data=tl.rates3[grepl("gi2",tl.rates3$genotype),],aes(y=rate,x=after.stratification_hr)  )
p.growthrate <- p.growthrate +
  geom_rect(data=rectangles,mapping=aes(xmin=night,xmax=day,ymin=bottom,ymax=top),fill="grey80")
p.growthrate <- p.growthrate +
  geom_smooth(mapping=aes(color=genotype),
              stat="smooth",
              fill="grey20",
              method="loess",span=.1,n=200, se = FALSE) # need to play with span and n
p.growthrate <- p.growthrate + theme(
  panel.background = element_rect(fill = "white")) +
  labs(x="Time (hr)",y="Growth rate (mm/hr)") +
   theme(axis.title = element_text(face="bold")) +
  theme(legend.position = "bottom", legend.title=element_blank(), legend.text = element_text(face = "italic"))
p.growthrate <- p.growthrate + coord_cartesian(xlim=c(100,max(tl.rates3$after.stratification_hr)), ylim=c(-.005,0.075))
#p.growthrate <- p.growthrate + coord_cartesian(ylim=c(-0.001,0.005))
p.growthrate <- p.growthrate + scale_color_manual(values = colors, labels = c("gi2"="gi-2", "pif3 gi2" = "gi2;pif3-1"))
p.growthrate
ggsave(p.growthrate,file="Maria_pif3gi2_timelapse_v3.png",dpi=300,height=2.7,width=3.3) 
```
